Wrinkling of a confined porous layer

### 2014-09-18 seminaire_Tanakalab.tex
Definition of biogels
Adapt intro

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,arrows.meta,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

@@

\begin{frame}{Application of biogels}
\begin{tabu}{cX[c]c}
\includegraphics[height=0.3\textheight]{cell_mech} &
\includegraphics[height=0.3\textheight]{electrophoresis} &
\includegraphics[height=0.3\textheight]{cosmetics} \\
Cell mechanics & Electrophoresis & Cosmetics\\
\end{tabu}
\begin{tabu}{X[c]X[c]}
\includegraphics[height=0.3\textheight]{bacterial_culture} &
\includegraphics[height=0.3\textheight]{food} \\
Bacterial culture & Food\\
\end{tabu}

@@

>>>>

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

@@

\begin{frame}{Application of biogels \textup{\normalsize(physical gels
$E_{bond}\gg k_B T$)}}
\begin{tabu}{cX[c]c}
\includegraphics[height=0.3\textheight]{cell_mech} &
\includegraphics[height=0.3\textheight]{electrophoresis} &
\includegraphics[height=0.3\textheight]{cosmetics} \\
Cell mechanics & Electrophoresis & Cosmetics\\
\end{tabu}
\begin{tabu}{X[c]X[c]}
\includegraphics[height=0.3\textheight]{bacterial_culture} &
\includegraphics[height=0.3\textheight]{food} \\
Bacterial culture & Food\\
\end{tabu}

@@

%\tikzset{external/force remake}
\begin{frame}{Over-acidification: Faster but weaker}
\tikzsetnextfilename{gelation_GDL}
\begin{tikzpicture}
\begin{axis}[
	height=0.8\textheight,
	width=\textwidth,
	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
	cycle list name=earthy,
	no marks,
	xmin=0, xmax=20,ymin=0,
	]
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
\end{axis}
\end{tikzpicture}

Caseins regain some solubility at low pH.
\end{frame}

### 2014-09-18 wokshop_wallslip.tex
Adapt intro

\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
\title[Yoghurt under stress]{Yoghurt under stress}
\author[M. Leocmach]{Mathieu Leocmach}
\date{1st June 2014}
\titlegraphic{
	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
	}

@@

\begin{frame}{Fundamental issues in biogels}
\begin{itemize}
\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
\end{itemize}
\begin{tabu}{X[c]X[c]}
\includegraphics[height=6\baselineskip]{Bonn_fracture} &
\includegraphics[height=6\baselineskip]{Villi_sq}
\end{tabu}
\begin{block}{Our questions}
\begin{itemize}
\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
\item What is the role of the solvent in mechanical instabilities?
\end{itemize}
\end{block}
\end{frame}

@@

\begin{columns}
\column{0.5\textwidth}
\tikzsetnextfilename{gelation}
\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group name=g, group size=1 by 2,
			xticklabels at=edge bottom,
			vertical sep=0,
			},
		%xmode=log,
		%xmin=1e2,xmax=3e5,
		xmin=0, xmax=20,
		scale only axis,
		width=\textwidth-4em,
		height=0.3\textwidth,
		extra tick style={grid=major},%
		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
		]
	\nextgroupplot[
		ymin=0, ymax=7, ylabel={pH},
		extra y ticks={4.6}, extra y tick labels={},%
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y190_cas4_gdl1.pH};
	\draw[help lines] (axis cs:0,4.6) -- (axis cs:3.88,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:3.88,0);

	\nextgroupplot[
		xlabel={time (\si{\hour})},
		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
		extra x ticks={17}, extra x tick labels={},
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y22.prise};
	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL1_Y22.prise};
	\begin{scope}[font=\scriptsize]
		%\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
		\node[anchor=base,rotate=80] at (axis cs:2.5,400) {gelation};
	\end{scope}
	\end{groupplot}
	\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
\end{tikzpicture}

@@

\begin{scriptsize}
\textit{Kaláb (1983),\linebreak Roefs \& van Vliet (1990),\linebreak Lucey \& Singh (1998)}
\end{scriptsize}
\end{columns}
\end{frame}
\tikzset{external/force remake=false}

\begin{frame}{Linear rheology: Power law (visco)elastic solid}
\[\text{stress}\rightarrow \sigma = G \gamma\leftarrow\text{strain}\]
\[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
\tikzsetnextfilename{power_law_rheology}
\begin{tikzpicture}
\begin{loglogaxis}[
	height=0.8\textheight,
	width=\textwidth,
	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{loss} moduli (\si{\pascal})},
	domain={6e-2:70},
	]
	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};	
\end{loglogaxis}
\end{tikzpicture}
\end{frame}

>>>>

\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
\title[Yoghurt wrinkling]{Wrinkling of a confined porous layer}
\author[M. Leocmach]{Mathieu Leocmach}
\date{24\textsuperscript{th} September 2014}
\titlegraphic{
	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
	}

@@

\begin{frame}{Fundamental issues in biogels}
\begin{itemize}
\item Gel stability and solvent expulsion (syn\ae{}resis) \textit{\scriptsize Bartlett et al. PRE 2012}
\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
\end{itemize}
\begin{tabu}{X[c]X[c]}
\includegraphics[height=8\baselineskip]{faisselle} &
\includegraphics[height=8\baselineskip]{Villi_sq}
\end{tabu}
\begin{block}{Our questions}
\begin{itemize}
\item What is the role of boundary conditions?
\item What is the role of the solvent in mechanical instabilities?
\end{itemize}
\end{block}
\end{frame}

@@

\begin{columns}
\column{0.5\textwidth}
\tikzsetnextfilename{gelation}
\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group name=g, group size=1 by 2,
			xticklabels at=edge bottom,
			vertical sep=0,
			},
		%xmode=log,
		%xmin=1e2,xmax=3e5,
		xmin=0, xmax=4,
		scale only axis,
		width=\textwidth-4em,
		height=0.3\textwidth,
		extra tick style={grid=major},%
		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
		]
	\nextgroupplot[
		ymin=0, ymax=7, ylabel={pH},
		extra y ticks={4.6}, extra y tick labels={},%
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y189_cas4_gdl4.pH};
	\draw[help lines] (axis cs:0,4.6) -- (axis cs:0.26,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:0.26,0);

	\nextgroupplot[
		xlabel={time (\si{\hour})},
		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
		extra x ticks={17}, extra x tick labels={},
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise};
	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL4_Y271.prise};
	\begin{scope}[font=\scriptsize]
		%\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
		\node[anchor=base,rotate=80] at (axis cs:0.2,300) {gelation};
	\end{scope}
	\end{groupplot}
	%\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
\end{tikzpicture}

@@

\begin{scriptsize}
\textit{Kaláb (1983),\linebreak Roefs \& van Vliet (1990),\linebreak Lucey \& Singh (1998)}
\end{scriptsize}
\end{columns}
\end{frame}
%\tikzset{external/force remake=false}

%\tikzset{external/force remake}
\begin{frame}{Over-acidification: Faster but weaker}
\tikzsetnextfilename{gelation_GDL}
\begin{tikzpicture}
\begin{axis}[
	height=0.8\textheight,
	width=\textwidth,
	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
	cycle list name=earthy,
	no marks,
	xmin=0, xmax=20,ymin=0,
	]
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
\end{axis}
\end{tikzpicture}

Caseins regain some solubility at low pH.
\end{frame}
